On the adhesivity of EGGS

### 2025-02-26 calco/main.tex
minima
correzioni

This  section briefly recalls \emph{$\mathcal{M}$-adhesive categories} \cite{azzi2019essence,ehrig2012,ehrig2014adhesive,lack2005adhesive,heindel2009category}. 
%
%\begin{notation}
Given a category $\X$ we do not distinguish notationally between $\X$ and its class of objects, so
``$X\in \X$'' means that $X$ is an object of $\X$. We let $\mor(\X)$, $\mon(\X)$ and $\reg(\X)$ denote the class of all arrows, monos and regular monos of $\X$, respectively.  Given an object $X$, we  denote by $?_X$ the unique arrow from an initial object into $X$ and by $!_X$ that  unique arrow from $X$ into a terminal one.
%\end{notation}

@@

We will briefly list some examples of $\mathcal{M}$-adhesive categories.

@@

\begin{restatable}{proposition}{epic}\label{prop:reg_epi_coeq}
    Let $e\colon X \to Y$ be a regular epi in a category $\X$ with a kernel pair $(K_e, \pi^1_e, \pi^2_e)$. Then, $e$ is the coequalizer of $\pi^1_e$ and $\pi^2_e$.
\end{restatable}
%[Proof in \Cref{epic-proof}]

\begin{restatable}{corollary}{natepi}\label{cor:reg_epi_components_reg_epi_nat_trans}
    Let $\X$ be a category with pullbacks and $\phi\colon F \dot\to G$ a natural transformation between functors $F, G: \D \rightrightarrows \X$. If $\phi_d$ is a regular epi for every $d$ in $\D$, then $\phi$ is a regular epi.
\end{restatable}
%[Proof in \Cref{natepi-proof}]

@@

In this section we briefly recall the notion \emph{$\X$-hypergraph}. It is necessary to have a monad $(-)^\star\colon \Set\to \Set$, also known as \emph{list monad},
sending a set to the free monoid on it \cite{sakarovitch2009elements,Wadler95} and playing a role analogous to the usual \emph{Kleene star}.
We recall some of its proprieties.

@@

\begin{remark}\label[remark]{rem:length}Notice that $1^\star$ can be canonically identified with $\mathbb{N}$, thus for every set $X$ the arrow $!_\X\colon X\to 1$ induces a \emph{length function} $!^\star_X\colon X^\star \to \mathbb{N}$, which sends a word to its length.
\end{remark}

@@

\todo{forse possiamo anche cancellare tutti gli esempi sotto}
\todo{FG ne lascerei uno, ripreso magari dalla intro e presentato come string diagram}

We can represent hypergraphs graphically. We will use dots to denote nodes and squares to denote hyperedges, the name of a node or of an hyperedge will be put near the corresponding dot or square. Sources and targets are represented by lines between dots and squares: the lines from the sources of an hyperedge will have an arrowhead in the middle pointing towards the hyperedge, while the lines to the targets will have arrowheads pointing to the target nodes.  We will decorate the arrow corresponding to the $i^{th}$ letter of a target or a source with a label $i$.

\begin{example}Take $V_{\mathcal{G}}$ to be be $\{v_1, v_2, v_3, v_4, v_5\}$ and $E_{\mathcal{G}}$ to be $\{h_1, h_2, h_3\}$. Sources and targets are given by:
	\[\begin{matrix}
		s_{\mathcal{G}}(h_1)\colon 2\to V_{\mathcal{G}}  & \begin{matrix}
			0 \mapsto v_1\\
			1\mapsto v_2
		\end{matrix} && s_{\mathcal{G}}(h_2)\colon 2\to V_{\mathcal{G}} & \begin{matrix}
			0 \mapsto v_3\\
			1\mapsto v_4 
		\end{matrix} && s_{\mathcal{G}}(h_3)\colon 1\to V_{\mathcal{G}} & 
		0 \mapsto v_5\\
		t_{\mathcal{G}}(h_1)\colon 2\to V_{\mathcal{G}} & \begin{matrix}
			0 \mapsto v_3\\
			1\mapsto v_4
		\end{matrix} && t_{\mathcal{G}}(h_2)\colon 2\to V_{\mathcal{G}} & 0\mapsto v_5 && t_{\mathcal{G}}(h_3)\colon 0\to V_{\mathcal{G}} &  t_{\mathcal{G}}(h_3)=?_{ V_{\mathcal{G}}} 
	\end{matrix}\]

@@

			\draw[->-=.5](1.75,-0.9)..controls(1.8,-1.3)and(2.5,-1.5)..(D)node[pos=0.5, below,font=\fontsize{7}{0}\selectfont]{$2$};
			\draw[->-=.5] (1.75,-0.6)..controls(1.8,-0.25)and(2.5,0)..(C) node[pos=0.5, above,font=\fontsize{7}{0}\selectfont]{$1$};
		\end{tikzpicture}
	\end{center}
\end{example}
\begin{example}\label[example]{exa_2} Let $V_{\mathcal{G}}$ be as in the previous example and $E_{\mathcal{G}}=\{h_1, h_2, h_3\}$.	Then we define
	\[\begin{matrix}
		s_{\mathcal{G}}(h_1)\colon 0\to V_{\mathcal{G}}  & s_{\mathcal{G}}(h_1)=?_{V_\mathcal{G}} && s_{\mathcal{G}}(h_2)\colon 2\to V_{\mathcal{G}} & \begin{matrix}
			0 \mapsto v_1\\
			1\mapsto v_2
		\end{matrix}&& s_{\mathcal{G}}(h_3)\colon 2\to V_{\mathcal{G}} & \begin{matrix} 
			0 \mapsto v_1\\
			1\mapsto v_4	
		\end{matrix}\\
		t_{\mathcal{G}}(h_1)\colon 1\to V_{\mathcal{G}} & 
		0 \mapsto v_1 && 		t_{\mathcal{G}}(h_2)\colon 1\to V_{\mathcal{G}} & 0\mapsto v_3 &&   t_{\mathcal{G}}(h_3)\colon 1\to V_{\mathcal{G}} & 1\mapsto v_5
	\end{matrix}\]

@@

\subsubsection{$\hyp$ as a category of functors}

@@

\begin{proposition}
	$\hyp$ is equivalent to the category $\Set^{\catname{H}}$.
\end{proposition}

@@

@@

\begin{remark}\label[remark]{rem:label}	
Let $\mathcal{H}=(E, V, s, t)$ be an hypergraph, by definition we know that $U_{\hyp}(\mathcal{G}^{\Sigma})$ is the terminal object $1$, so an arrow $\mathcal{H}\rightarrow \mathcal{G}^{\Sigma}$, is determined by a morphism $h\colon E_\mathcal{H}\to O_\Sigma$  making the two squares below commute (cfr.~ \Cref{rem:length}).
	\[\xymatrix{E_{\mathcal{H}} \ar[r]^{h} \ar[d]_{s_{\mathcal{H}}}& O_\Sigma \ar[d]^{\ari_\Sigma}& E_{\mathcal{H}} \ar[r]^{h} \ar[d]_{t_{\mathcal{H}}} & O_\Sigma \ar[d]^{\delta_1}\\ V^\star_{\mathcal{{H}}} \ar[r]_{\lgh_{V_{\mathcal{H}}}}& \mathbb{N} & V^\star_{\mathcal{{H}}} \ar[r]_{\lgh_{V_{\mathcal{H}}}} & \mathbb{N}}\]
%	
Let $v_n\colon V^n_\mathcal{H}\to  V^\star_{\mathcal{H}}$ be a coprojection.  The second diagram above entails that $t_{\mathcal{H}}$ factors via the inclusion $v_1\colon V_{\mathcal{H}}\to V^{\star}$ of words of length $1$, i.e. $t_{\mathcal{H}}=v_1\circ \tau_{\mathcal{H}}$ for some $\tau_{\mathcal{H}}\colon E_{\mathcal{H}}\to \tau_{\mathcal{H}}$.
\end{remark}

$\hyp_{\Sigma}$, has a forgetful functor $U_{\Sigma}\colon \hyp_{\Sigma}\to \X$ which sends $(h,k)\colon \mathcal{H}\to \mathcal{G}^{\Sigma}$ to $U_{\X}(\mathcal{H}$). Now, $U_{\X}(\mathcal{G}^{\Sigma})=1$ thus, for every object $X$, there is only one arrow $X\to U_{\X}(\mathcal{G}^{\Sigma})$. Define $\Delta_{\Sigma}(X)\colon \Delta_{\X}(X)\to \mathcal{G}^{\Sigma}$ as the transpose of this arrow. Explicitly, $\Delta_{\X}(X)=(0, X, ?_{X^\star}, ?_{X^\star})$ and $\Delta_\Sigma(X)$ is simply $(?_{O_\Sigma}, !_{X})$.

\begin{proposition} $U_\Sigma$
	has a left adjoint $\Delta_\Sigma$.
\end{proposition}
\begin{proof}Let $(h, !_{V_\mathcal{H}})\colon \mathcal{H}\to \mathcal{G}^{\Sigma}$ be an object of $\hyp_{\Sigma}$, and suppose that there exists $f\colon X\to U_{\Sigma}(\mathcal{H})$. Since, $U_{\Sigma}(\mathcal{H})=U_{\X}(\mathcal{H})$ and the identity is the unit of $\Delta_\hyp \dashv U_{\hyp}$, we get a morphism $(?_{E_{\mathcal{H}}},f)\colon \Delta_{\X}(X)\to \mathcal{H}$ of $\hyp$. And then the thesis follows since we have
	\[
	(h, !_{V_{\mathcal{{H}}}})\circ (?_{E_{\mathcal{H}}}, f)=(h\circ ?_{E_{\mathcal{H}}}, !_{V_{\mathcal{{H}}}}\circ f)=(?_{0_{\Sigma}}, !_{X})=\Delta_{\hyp}(X)\]
%and the thesis follow.   
\end{proof}

@@

We will extend our graphical notation of hypergraphs to labeled ones putting the label of an hyperedge $h$ inside its corresponding square.
\begin{example}\label[example]{lab_1}
	The simplest example is given by the identity $\id{\mathcal{G}^\Sigma}\colon \mathcal{G}^\Sigma\rightarrow \mathcal{G}^{\Sigma}$. If $\Sigma$ is the signature of groups $\Sigma_G$we get 
	\begin{center}
		\begin{tikzpicture}
			\node[circle,fill=black,inner sep=0pt,minimum size=6pt,label=above:{$\heartsuit$}] (V) at (0,0) {};
			\node(E)at(-2, 0.4){$e$};
			\node(M)at(0, 2.15){$\cdot$};
			\node(I)at(2, 0.75){$(-)^{-1}$};

@@

\begin{example}\label[example]{lab_2}
	Take again $\Sigma_G$ the signature of groups, then the hypergraph $\mathcal{G}$ of \Cref{exa_2} can be labeled defining
	\begin{align*}
		e=f(h_1) \quad \cdot=f(h_2)\quad \cdot=f(h_3)
	\end{align*}
	In this case we get the following picture

@@

\input{Hypergraphs_with_equivalences}
\input{Labeled_hyp_with_eq}

>>>>

This  section briefly recalls \emph{$\mathcal{M}$-adhesive categories} \cite{azzi2019essence,ehrig2012,ehrig2014adhesive,lack2005adhesive,heindel2009category}. 
%
%\begin{notation}
Given a category $\X$ we do not distinguish notationally between $\X$ and its class of objects, so
``$X\in \X$'' means that $X$ is an object of $\X$. We let $\mor(\X)$, $\mon(\X)$ and $\reg(\X)$ denote the class of all arrows, monos and regular monos of $\X$, respectively.  Given an object $X$, we  denote by $?_X$ the unique arrow from an initial object into $X$ and by $!_X$ that  unique arrow from $X$ into a terminal one. We will also use the notation $e\colon X\eto Y$ to denote that an arrow $e\colon X\to Y$ is a regular epi. 
%\end{notation}

@@

We briefly list some examples of $\mathcal{M}$-adhesive categories.

@@

\begin{restatable}{proposition}{epic}\label{prop:reg_epi_coeq}
    Let $e\colon X \eto Y$ be a regular epi in a category $\X$ with a kernel pair $(K_e, \pi^1_e, \pi^2_e)$. Then, $e$ is the coequalizer of $\pi^1_e$ and $\pi^2_e$.
\end{restatable}
%[Proof in \Cref{epic-proof}]

\begin{restatable}{corollary}{natepi}\label{cor:reg_epi_components_reg_epi_nat_trans}
    Let $\X$ be a category with pullbacks and $\phi\colon F \to G$ a natural transformation between functors $F, G: \D \rightrightarrows \X$. If $\phi_d$ is a regular epi for every $d$, then $\phi$ is a regular epi.
\end{restatable}
%[Proof in \Cref{natepi-proof}]

@@

In this section we briefly recall the notion of \emph{hypergraph}. In order to so, a pivotal role is played by the \emph{Kleen star} monad $(-)^\star\colon \Set\to \Set$, also known as \emph{list monad},
sending a set to the free monoid on it \cite{sakarovitch2009elements,Wadler95}.
We recall some of its proprieties.

@@

\begin{remark}\label[remark]{rem:length}Notice that $1^\star$ can be canonically identified with $\mathbb{N}$, thus for every set $X$ the arrow $!_\X\colon X\to 1$ induces a \emph{length function} $\lgh_{X}\colon X^\star \to \mathbb{N}$, which sends a word to its length.
\end{remark}

@@

\todo{FG ne lascerei uno, ripreso magari dalla intro e presentato come string diagram}

We can represent hypergraphs graphically. We will use dots to denote nodes and squares to denote hyperedges, the name of a node or of an hyperedge will be put near the corresponding dot or square. Sources and targets are represented by lines between dots and squares: the lines from the sources of an hyperedge will have an arrowhead in the middle pointing towards the hyperedge, while the lines to the targets will have arrowheads pointing to the target nodes.  We will decorate the arrow corresponding to the $i^{th}$ letter of a target or a source with a label $i$.
\todo{Ho lasciato un unico esempio (che è anche un term graph): occorre riscrivere questo paragrafo una volta che abbiamo deciso come rappresentare i grafi}
\iffalse 
\begin{example}Take $V_{\mathcal{G}}$ to be be $\{v_1, v_2, v_3, v_4, v_5\}$ and $E_{\mathcal{G}}$ to be $\{h_1, h_2, h_3\}$. Sources and targets are given by:
	\[\begin{matrix}
		s_{\mathcal{G}}(h_1)\colon 2\to V_{\mathcal{G}}  & \begin{matrix}
			0 \mapsto v_1\\
			1\mapsto v_2
		\end{matrix} && s_{\mathcal{G}}(h_2)\colon 2\to V_{\mathcal{G}} & \begin{matrix}
			0 \mapsto v_3\\
			1\mapsto v_4 
		\end{matrix} && s_{\mathcal{G}}(h_3)\colon 1\to V_{\mathcal{G}} & 
		0 \mapsto v_5\\
		t_{\mathcal{G}}(h_1)\colon 2\to V_{\mathcal{G}} & \begin{matrix}
			0 \mapsto v_3\\
			1\mapsto v_4
		\end{matrix} && t_{\mathcal{G}}(h_2)\colon 2\to V_{\mathcal{G}} & 0\mapsto v_5 && t_{\mathcal{G}}(h_3)\colon 0\to V_{\mathcal{G}} &  t_{\mathcal{G}}(h_3)=?_{ V_{\mathcal{G}}} 
	\end{matrix}\]

@@

			\draw[->-=.5](1.75,-0.9)..controls(1.8,-1.3)and(2.5,-1.5)..(D)node[pos=0.5, below,font=\fontsize{7}{0}\selectfont]{$2$};
			\draw[->-=.5] (1.75,-0.6)..controls(1.8,-0.25)and(2.5,0)..(C) node[pos=0.5, above,font=\fontsize{7}{0}\selectfont]{$1$};
		\end{tikzpicture}
	\end{center}
\end{example}
\fi

\begin{example}\label[example]{exa_2} Let $V_{\mathcal{G}}$ be as in the previous example and $E_{\mathcal{G}}=\{h_1, h_2, h_3\}$.	Then we define
	\[\begin{matrix}
		s_{\mathcal{G}}(h_1)\colon 0\to V_{\mathcal{G}}  & s_{\mathcal{G}}(h_1)=?_{V_\mathcal{G}} && s_{\mathcal{G}}(h_2)\colon 2\to V_{\mathcal{G}} & \begin{matrix}
			0 \mapsto v_1\\
			1\mapsto v_2
		\end{matrix}&& s_{\mathcal{G}}(h_3)\colon 2\to V_{\mathcal{G}} & \begin{matrix} 
			0 \mapsto v_1\\
			1\mapsto v_4	
		\end{matrix}\\
		t_{\mathcal{G}}(h_1)\colon 1\to V_{\mathcal{G}} & 
		0 \mapsto v_1 && 		t_{\mathcal{G}}(h_2)\colon 1\to V_{\mathcal{G}} & 0\mapsto v_3 &&   t_{\mathcal{G}}(h_3)\colon 1\to V_{\mathcal{G}} & 1\mapsto v_5
	\end{matrix}\]

@@

\begin{remark}\label{rem:functor}
	It is worth to point out, as first noted in \cite{bonchi2022string}, that $\hyp$ is actually equivalent to a category of functors.

	Indeed, consider the category $\catname{H}$ in which the set of objects is given by $ (\mathbb{N}\times \mathbb{N}) \cup \{\bullet\}$ and arrows are given by the identities $\id{k,l}$, $\id{\bullet}$ and exactly $k+l$ arrows $f_i\colon (k,l)\rightarrow \bullet$, where $i$ ranges from $0$ to $k+l-1$.

	It is quite straightforward to see that functors $\catname{H}\to \Set$ corresponds exactly to hypergaphs: nodes correspond to the image of $\bullet$ while the set of hyperedges with source of length $k$ and target of length $l$ corresponds to the image of $(k,l)$ (see \cite{CastelnovoGM24}).
\end{remark}

In particular, \Cref{rem:functor} entails the following result.

\begin{proposition}\label{prop:cocomp}
	$\hyp$ has all limits and colimits.
\end{proposition}

\iffalse
\subsubsection{$\hyp$ as a category of functors}

@@

\begin{proposition}
	$\hyp$ is equivalent to the category $\Set^{\catname{H}}$.
\end{proposition}
\fi

@@

\todo{Decidere se tenere questo esempio}

@@

\begin{remark}\label[remark]{rem:label}	
Let $\mathcal{H}=(E, V, s, t)$ be an hypergraph, by definition we know that $U_{\hyp}(\mathcal{G}^{\Sigma})$ is the terminal object $1$, so an arrow $\mathcal{H}\rightarrow \mathcal{G}^{\Sigma}$, is determined by a morphism $h\colon E_\mathcal{H}\to O_\Sigma$  making the two squares below commute (cfr.~ \Cref{rem:length}).
	\[\xymatrix{E_{\mathcal{H}} \ar[r]^{h} \ar[d]_{s_{\mathcal{H}}}& O_\Sigma \ar[d]^{\ari_\Sigma}& E_{\mathcal{H}} \ar[r]^{h} \ar[d]_{t_{\mathcal{H}}} & O_\Sigma \ar[d]^{\delta_1}\\ V^\star_{\mathcal{{H}}} \ar[r]_{\lgh_{V_{\mathcal{H}}}}& \mathbb{N} & V^\star_{\mathcal{{H}}} \ar[r]_{\lgh_{V_{\mathcal{H}}}} & \mathbb{N}}\]
%	
Let $v_n\colon V^n_\mathcal{H}\to  V^\star_{\mathcal{H}}$ be a coprojection.  The second diagram above entails that $t_{\mathcal{H}}$ factors via the inclusion $v_1\colon V_{\mathcal{H}}\to V^{\star}$ of words of length $1$, i.e. $t_{\mathcal{H}}=v_1\circ \tau_{\mathcal{H}}$ for some $\tau_{\mathcal{H}}\colon E_{\mathcal{H}}\to V_{\mathcal{H}}$.
\end{remark}

$\hyp_{\Sigma}$, has a forgetful functor $U_{\Sigma}\colon \hyp_{\Sigma}\to \X$ which sends $(h,k)\colon \mathcal{H}\to \mathcal{G}^{\Sigma}$ to $U_{\X}(\mathcal{H}$). Now, $U_{\hyp}(\mathcal{G}^{\Sigma})=1$ thus, for every set $X$ we can define $\Delta_{\Sigma}(X)\colon \Delta_{\X}(X)\to \mathcal{G}^{\Sigma}$  as $(?_{O_\Sigma}, !_{X})$. It is straightforward to see that in this way we get a left adjoint to $U_\Sigma$.

\begin{proposition} $U_\Sigma$
	has a left adjoint $\Delta_\Sigma$.
\end{proposition}
\iffalse 
\begin{proof}Let $(h, !_{V_\mathcal{H}})\colon \mathcal{H}\to \mathcal{G}^{\Sigma}$ be an object of $\hyp_{\Sigma}$, and suppose that there exists $f\colon X\to U_{\Sigma}(\mathcal{H})$. Since, $U_{\Sigma}(\mathcal{H})=U_{\X}(\mathcal{H})$ and the identity is the unit of $\Delta_\hyp \dashv U_{\hyp}$, we get a morphism $(?_{E_{\mathcal{H}}},f)\colon \Delta_{\X}(X)\to \mathcal{H}$ of $\hyp$. And then the thesis follows since we have
	\[
	(h, !_{V_{\mathcal{{H}}}})\circ (?_{E_{\mathcal{H}}}, f)=(h\circ ?_{E_{\mathcal{H}}}, !_{V_{\mathcal{{H}}}}\circ f)=(?_{0_{\Sigma}}, !_{X})=\Delta_{\hyp}(X)\]
%and the thesis follow.   
\end{proof}
\fi

@@

We will extend our graphical notation of hypergraphs to labeled ones putting the label of an hyperedge $h$ inside its corresponding square.
\begin{example}\label[example]{lab_1}
	The simplest example is given by the identity $\id{\mathcal{G}^\Sigma}\colon \mathcal{G}^\Sigma\rightarrow \mathcal{G}^{\Sigma}$. If $\Sigma$ is the signature of groups $\Sigma_G$ we get 
	\begin{center}
		\begin{tikzpicture}
			\node[circle,fill=black,inner sep=0pt,minimum size=6pt,label=above:{$\heartsuit$}] (V) at (0,0) {};
			\node(E)at(-2, 0.4){$e$};
			\node(M)at(0, 2.15){$\cdot$};
			\node(I)at(2, 0.75){$(-)^{-1}$};

@@

\todo{Decidere se tenere questo esempio (io direi di no)}
\begin{example}\label[example]{lab_2}
	Take again $\Sigma_G$ the signature of groups, then the hypergraph $\mathcal{G}$ of \Cref{exa_2} can be labeled defining
	\begin{align*}
		e=f(h_1) \quad \cdot=f(h_2)\quad \cdot=f(h_3)
	\end{align*}
	In this case we get the following picture

@@

\section{Adding equivalences to hypergraphical structures}
\label{hypereq}
\todo{A very nice instroduction}

\subsection{Hypergraphs with equivalence}

Let us start with the case of general hypergraphs.

\begin{definition}
	A \emph{hypergraph with equivalence} $\mathcal{G} = (E_\mathcal{G}, V_{\mathcal{G}}, Q_\mathcal{G}, s_\mathcal{G}, t_\mathcal{G}, q_\mathcal{G})$ is a 6-tuple such that $\mathcal{G} = (E_\mathcal{G}, V_{\mathcal{G}}, s_\mathcal{G}, t_\mathcal{G})$ is a hypergraph, $Q_\mathcal{G}$ is a set and $q_{\mathcal{G}}: V_{\mathcal{G}}\eto Q_{\mathcal{G}}$ is a surjection called \emph{quotient map}. 
	%
	A morphism $h\colon \mathcal{G\to H}$ is a triple $(h_E, h_V, h_C)$ such that the following diagrams commute
	\[\xymatrix{
		{E_\mathcal{G}}\ar[r]^{s_\mathcal{G}}\ar[d]_{h_E} & {V_{\mathcal{G}}^\star}\ar[d]^{h_V^\star} & {E_\mathcal{G}}\ar[r]^{t_\mathcal{G}}\ar[d]_{h_E} & {V_{\mathcal{G}^\star}}\ar[d]^{h_V^\star} & {V_\mathcal{G}}\ar@{>>}[r]^{q_\mathcal{G}}\ar[d]_{h_V} & {Q_{\mathcal{G}}} \ar[d]^{h_C} \\
		{E_\mathcal{H}}\ar[r]_{s_\mathcal{H}} & {V_{\mathcal{H}}^\star}	& {E_\mathcal{H}}\ar[r]_{t_\mathcal{H}} & {V_{\mathcal{H}^\star}}& {V_\mathcal{H}}\ar@{>>}[r]_{q_\mathcal{H}} & {Q_\mathcal{H}}
	}\]
	The category of hypergraphs with equivalences and their morphisms is denoted $\EqHyp$.

\end{definition}

\begin{remark}
	Notice that in $\Set$ the classes of surjections, epis and regular epis coincide.
\end{remark}

\begin{remark}\label{rem:eqhyp_morphs}
	Morphisms of hypergraphs with equivalences are uniquely determined by the first two components. That is, if $h_1 = (h_E, h_V, f)$ and $h_2 = (h_E, h_V, g)$ are two morphisms $\mathcal{G} \rightrightarrows \mathcal{H}$, then we have
	\[f \circ q_\mathcal{G} = q_\mathcal{H}\circ h_V =g\circ q_\mathcal{G}\]

	Since $q_\mathcal{G}$ is epi, we obtain $f = g$.
\end{remark}

Forgetting the quotient part yields a functor $T\colon \EqHyp \to \hyp$ which sends an hypergraph with equivalence $(E_\mathcal{G}, V_{\mathcal{G}}, C_\mathcal{G}, s_\mathcal{G}, t_\mathcal{G}, q_\mathcal{G})$ to $(E_{\mathcal{G}}, V_{\mathcal{G}}, s_\mathcal{G}, t_{\mathcal{G}})$.   We can now explore some of its properties to deduce some information on the structure of $\EqHyp$.

\begin{restatable}{proposition}{fhyp}  The following are true:
	\begin{enumerate}
		\item$T$ is faithful;
		\item $T$ has a left adjoint;
		\item $T$ has a right adjoint.
	\end{enumerate}
\end{restatable}
A proof is provided in \Cref{prop:forghyp}.

\begin{corollary}\label{cor:limcolim}
	The functor $T$ preserves limits and colimits.
\end{corollary}

Combining together the previous corollary with \Cref{rem:eqhyp_morphs} we get the following characterization of monos in $\EqHyp$.

\begin{corollary}\label{cor:mono1}
	An arrow $h = (h_E, h_V, h_C): \mathcal{G \to H}$ in $\EqHyp$ is mono if and only if $(h_E, h_V)$ is a mono in $\hyp$.
\end{corollary}

\commentato{
	\todo{questo non serve a nulla e in più è pure scritto sbagliato}
$\EqHyp$ has a forgetful functor $U_{\EqHyp}:\EqHyp \to \Set$, which sends each $\mathcal{G} = (E_\mathcal{G}, V_{\mathcal{G}}, C_\mathcal{G}, s_\mathcal{G}, t_\mathcal{G}, q_\mathcal{G})$ into $V_\mathcal{G}$, and each $h = (h_E, h_V, h_C)$ onto $h_V$.

\begin{corollary}\label{cor:ladj}
	$U_\EqHyp$ has a left adjoint $\Delta_{\EqHyp}: \Set \to \EqHyp$.
\end{corollary}

\begin{proof}
	For each set $X$, define $\Delta_\EqHyp(X):= (\emptyset, X, \{\bullet\}, ?_X, ?_X, !_X)$. Consider now $h: \Delta_{\EqHyp}(X) \to \mathcal{H}$.
	\[\xymatrix@C=2.3cm{
		\Delta_{\EqHyp}(X) \ar@{.>}[d]_{\Delta_\EqHyp(f)} \ar[dr]^{h} & \\
		\Delta_\EqHyp(U_{\EqHyp}(\mathcal{H})) \ar[r]_{\epsilon_{\mathcal{H}}} & \mathcal{H}
	}\]
	Where $\Delta_\EqHyp(U_{\EqHyp}(\mathcal{H})) = (\emptyset, V_\mathcal{H}, \{\bullet\}, ?_{V_\mathcal{H}}, ?_{V_\mathcal{H}}, !_{V_\mathcal{H}})$ and $\epsilon_{\mathcal{H}} = (?_{E_\mathcal{H}}, \id{V_\mathcal{H}}, g)$.
	Note that, since $\Delta_{\EqHyp}(X)$ has the empty set as object of edges, $h_E = ?_{E_\mathcal{H}}$, then, the unique arrow that fits in the diagram is $\Delta_{\EqHyp}(f) = (?_{E_\mathcal{H}}, h_V, \id{\{\bullet\}})$.
\end{proof}}

Let us notice that there $K: \EqHyp \to \Set$ sending $(E, V, C, s, t, q)$ to $C$, and a morphism $(h_E, h_V, h_C)$ to $h_C$. We are now going to exploit it to compute limits and colimits in $\EqHyp$. In order to do so. A full proof of the following lemma can be found in \Cref{proof:comp}.

\noindent
\parbox{11cm}{
\begin{restatable}{lemma}{comp}\label{prop:eqhyp_complete}
Let $F\colon \D \to \EqHyp$ be a diagram, and let $(\mathcal{E}_d, V_d, Q_d, s_d, t_d, q_d)$ be the image of an object $d$. Then the following hold true:
\begin{enumerate}
		\item $F$ has a colimit which is preserved by $K$;
	\item let $(L, \{l_d\}_{d\in \D})$ be a limiting cocone for $K \circ F$, then $F$ has a limit $(E, V, Q, s, t, q)$ and there is a mono $m\colon Q\mto L$ such that the diagram on the right commutes for every $d\in \D$, where $((E, V), \{(\pi^d_E, \pi^d_V)\}_{d\in \D})$ is limiting for $T\circ F$.
\end{enumerate}
\end{restatable}}\hfill 
\parbox{2cm}{\xymatrix{V \ar@{>>}[d]_{q} \ar[r]^{\pi^d_V}& V_d \ar@{>>}[dd]^{q_d}\\Q \ar@{>->}[d]_{m}&\\ L \ar[r]_{l_d} & Q_d}}

\begin{corollary}\label{cor:mono2}
	If $h = (h_E, h_V, h_C): \mathcal{G\to H}$ is a regular mono in $\EqHyp$, then $h_E, h_V$ and $h_C$ are all monos.
\end{corollary}
\todo{Questa è da rifare completamente}
\begin{proof}
	If $h$ is mono, from \Cref{cor:mono1} we have that $h_E$ and $h_V$ are monos. Suppose now $f, g: \mathcal{H\rightrightarrows K}$ be the arrows equalized by $h$. Then, we have:
	\begin{align*} f_C \circ h_C \circ q_\mathcal{G} &= f_C\circ q_\mathcal{H} \circ h_V\\&=q_\mathcal{K}\circ g_V \circ h_V\\&=q_\mathcal{K}\circ f_V \circ h_V\\&=g_C \circ h_C \circ q_\mathcal{G} \end{align*}
	Since $q_\mathcal{G}$ is epi, we have $f_C \circ h_C = g_C \circ h_C$, hence $h_C$ is an equalizer for $f_C$ and $g_C$, and thus a mono.
\end{proof}

\iffalse 
\begin{proposition}
	Let $h = (h_E, h_V, h_C): \mathcal{G \to H}$ be a regular mono in $\EqHyp$.
	Then, $h_E$ and $h_V$ are monos and $(K, \pi_1, \pi_2)$ is the kernel pair of $q_\mathcal{H}\circ h_V$ if and only if $(K, \pi_1, \pi_2)$ is the kernel pair of $q_\mathcal{G}$.
\end{proposition}

\begin{proof}
	By \Cref{cor:mono2}, we have that $h_E, h_V$ and $h_C$ are all monos.
	Hence, by \Cref{cor:kermono}, $(K, \pi_1, \pi_2)$ is the kernel pair of $q_\mathcal{G}$ if and only if it is the kernel pair also of $h_C \circ q_\mathcal{G}$, since $h_C$ is mono by hypothesis.
	The thesis follows from $h_C \circ q_\mathcal{G} = q_H \circ h_V$, and from the hypothesis of $h_E$ mono.
\end{proof}

\begin{remark}
	It is possible to restate the last proposition, by \Cref{ex:kernel_pairs_in_Set}, as 
	\begin{displayquote}
		\textit{$h_E$ and $h_V$ are mono and, for every $v, v'\in V_H$, $q_H(h_V(v))=q_H(h_V(v'))$ if and only if $q_G(v)=q_G(v')$}
	\end{displayquote}
	That is, a regular mono in $\EqHyp$ is a morphism that both reflects and preserves equivalences.
	% besides preserving them.
\end{remark}
\fi

Let us turn to another functor $\EqHyp \to \hyp$.

\begin{definition}
	The \emph{quotient functor} $Q:\EqHyp\to \hyp $ is defined as
	the one sending $(E, V, C, s, t, q)$ to $(E, C, q^{\star}\circ s, q^{\star}\circ t)$ and an arrow $(h_E, h_V, h_C)$ to $(h_E, h_C)$.
\end{definition}

\begin{remark}
	The action of the functor on a morphism of hypergraphs with equivalences gives a morphism of hypergraphs,
	in fact $q^{\star}_\mathcal{H} \circ s_\mathcal{H} \circ h_E = q^{\star}_\mathcal{H} \circ h_V^\star \circ s_\mathcal{G} = h_C^\star \circ q^{\star}_\mathcal{G} \circ s_\mathcal{G}$.
	The same is valid for $t_\mathcal{H}$ and $t_\mathcal{G}$. 
\end{remark}

\begin{lemma}\label{lemma:quot_funct_left_adj}
	$Q$ is a left adjoint.
\end{lemma}

\begin{proof}
	Let $R((A, B, s, t))$ be $(A, B, B, s, t, \id{B})$, so that $Q(R((A, B, s, t))) = (A, B, s, t)$. Now, suppose that $h = (h_E, h_V): Q((E, V, C, s', t', q)) \to (A, B, s, t)$  is an arrow in $\hyp$, and consider the triple $(h_E, h_V, h_V \circ q)$. Since $h$ is a morphism of $\hyp$, we have $h_V^{\star}\circ q^{\star}\circ s'= s\circ h_E$ and $h_V^{\star}\circ q^{\star}\circ t' = t\circ h_E$.
	Then we have the following squares
	\[\xymatrix{
		E \ar[r]^{h_E}\ar[d]_{s'} & A \ar[d]^{s} & E \ar[r]^{h_E}\ar[d]_{t'} & A \ar[d]^{t} & V \ar[r]^{h_V \circ q} \ar[d]_{q} & B \ar[d]^{\id{B}} \\
		V^{\star} \ar[r]_{h_V^\star\circ q^\star} & B^{\star} & V^{\star} \ar[r]_{h_V^\star\circ q^\star} & B^{\star} & C \ar[r]_{h_V} & B
	}\]

	\iffalse\[
	\begin{tikzcd}
		E \arrow[r, "{h_E}"] \arrow[d, "{s_G}"swap] & A \arrow[d, "s"] \\
		V \arrow[r, "{h_V \circ q}"swap] & B
	\end{tikzcd}
	\qquad
	\begin{tikzcd}
		E \arrow[r, "{h_E}"] \arrow[d, "{t_G}"swap] & A \arrow[d, "t"] \\
		V \arrow[r, "{h_V \circ q}"swap] & B
	\end{tikzcd}
	\qquad
	\begin{tikzcd}
		V \arrow[r, "{h_V\circ q}"] \arrow[d, "q" swap] & B \arrow[d, "{id_B}"] \\
		C \arrow[r, "{h_V}"swap] & B
	\end{tikzcd}
	\]\fi
	We have therefore found a morphism $(E, V, C, s', t', q) \to R((A, B, s, t))$ whose image through $Q$ fits in the diagram below
	\[\xymatrix{
		(A, B, s, t) \ar[r]^{(\id{A}, \id{B})} & (A, B, s, t)\\
		(E, C, q^\star\circ s', q^\star\circ t') \ar[u]^{Q(h_E, h_V \circ q, h_V)} \ar[ur]_{(h_E, h_V)}
	}\]   
	Such arrow is unique. Suppose $f = (f_E, f_V, f_C)$ to be another arrow with such property.
	Then, it must be $(\id A, \id B) \circ Q(f) = (f_E, f_C) = (h_E, h_C)$. Finally, $f_C = f_V \circ q = h_V \circ q$. 
\end{proof}
